---
title: "Assignment 1"
author: "Kyle Walker"
date: "9/18/2019"
output:
  pdf_document: default
  html_document:
    pdf_document: default
---
## Part I - R Basics

### 1. Calculate the following sums:
```{r warning=FALSE}
#S1
x <- c(1:2019)
sum(x)

#S2
sum(x^3)

#S3
sum(x^x)

s4 <- (c(1,-1)* x)

#S4
sum(s4^x)

#S5
sum(1/(x^2))

#S6
sum(1/x)

#S7
sum(1/(x^3))

#S8
sum(1/(s4))

```
### 2. The rnorm function generate random variables from normal distribution. Generate a sample of 1000 values from normal distribution with the mean 10 and standard deviation 1.
#### a. Calculate the mean and standard deviation of the sample.

```{r}
norm <- rnorm(1000, 10, 1)
mean(norm)
sd(norm)
```

#### b. Out of 1000 samples, how many do you think are that great than 10? Check your estimation.

```{r}
#Should expect 500
sum(norm > 10)
```

#### c. Use hist() function to show the histogram of the sample.

```{r}
hist(norm)

```

#### d. Estimate P(X > 1), where X follows N(2, 1)

```{r Warning=FALSE}
prob <- rnorm(1000, 2, 1)
#Probability that X > 1:
sum(prob > 1) / 1000
```

### 3. Consider an experiment of tossing a fair dice.
#### a. Use the sample (with replacement) function to generate a sample of 1000 values from the experiment.
```{r}
x <- c(1:6)
s <- sample(x, 1000, replace = TRUE)
```

#### b. Calculate the mean and standard deviation of the sample.
```{r}
mean(s)
sd(s)
```

#### c. How many times the 6 occured?
```{r}
sum(s == 6)
```

#### d. Use table function to show the frequency of the values.
```{r}
table(s)
```

#### e. Use prop.table(table()) to show the relative frequency of the values.
```{r}
prop.table(table(s))
```

#### f. Plot the frequency of the values.
```{r}
hist(s)
```

### 4. Consider an experiment of tossing a dice 3 times. Let X1, X2, and X3 be the number of tossing the first time, second time and third time, respectively. Use simulation to estimate the following probabilities:
#### a. P(X1 > X2 + X3)
```{r}
exper <- matrix(data=(sample(x, 3000, replace = TRUE)), nrow = 1000, ncol = 3)
sum(exper[,1] > (exper[,2] + exper[,3])) / 1000
```

#### b. P(X1^2 > X2^2 + X3^2)
```{r}
sum(exper[,1]^2 > (exper[,2]^2 + exper[,3]^2)) / 1000
```

### 5. Using simulation, estimate the probability of getting three tails in a row when tossing a coin 3 times.
#### Hint: one way is to generate a matrix with three columns where each rows is an observation of tossing a coin three times.

```{r}
# 1 is a head in this case
y <- c(0,1)

sim <- matrix(sample(y, 3000, replace= TRUE), nrow= 1000, ncol = 3)
sum((sim[,1] + sim[,2] + sim[,3]) == 0)/1000

```

### 6. (Extra Credits/Optional) Using simulation, estimate the probability of getting three tails in a row when tossing a coin 10 times.

```{r}
coin <- matrix(sample(c(0,1), 10000, replace= TRUE), nrow= 1000, ncol = 10)

count <- 0

consec <- function(x){
   sum(max(rle(x)$lengths >=3 & rle(x)$values == 0)) 
  }
  
sum(apply(coin, 1, consec)) / 1000




```

### 7. Central Limit Theorem (CLT). The CLT said that the mean of a sample of a distribution A (no matter what A is) follows normal distribution with the same mean as A. Following the below steps to confirm the CLT when A is uniform distribution.
#### • Generate 100 samples of uniform distibution from 0 to 1. Each sample has 1000 observations. Use the runif function to do this.
```{r}
uni <- matrix(runif(100000, 0, 1), nrow= 100, ncol = 1000)
```

#### • Compute the means of the 100 samples. Create vector x containing these means. Hint: You want to put all the samples in a matrix and use rowSums or colSums function.
```{r}
x <- rowSums(uni)/1000
```

#### • By CLT, x must follow normal distribution. Check this by plotting the histogram of x. Does it look like normal distribution? Use hist(x) to plot the histogram of x.
```{r}
hist(x)
```

#### • Increase the number (100 and 1000) to see if the distribution of x looks more like normal distribution.
```{r}
uni2 <- matrix(runif(1000000, 0, 1), nrow= 1000, ncol = 1000)

x <- rowSums(uni2) / 1000

hist(x)
```

#### • Try the same procedure with two other distributions for A.

```{r}
#Poisson
pois <- matrix(rpois(100000, 2), nrow= 100, ncol = 1000)

x <- rowSums(pois) / 1000

hist(x)

pois2 <- matrix(rpois(1000000, 2), nrow= 1000, ncol = 1000)

y <- rowSums(pois2) / 1000

hist(y)

#Binomial
binom <- matrix(rbinom(100000, 1, 0.5), nrow= 100, ncol = 1000)

x <- rowSums(binom) / 1000

hist(x)

binom2 <- matrix(rbinom(1000000,1, 0.5), nrow= 1000, ncol = 1000)

y <- rowSums(binom2) / 1000

hist(y)
```

### Part II - Working with Data
#### 7. Use read.csv function to read in the titanic dataset. You can find the dataset on Blackboard or at Kaggle.com. Use str function to see a summary of the data.
```{r}
titanic <- read.csv(file = "C:/Users/student/Documents/Senior Year/MATH 421/titanic.csv", header = T, sep= "," )
str(titanic)
```

#### 8. Use knitr::kable function to nicely print out the first 10 rows of the data in markdown.
```{r}
knitr::kable(head(titanic, 10))
```

#### 9. Use is.na function and sum function to count the total number of missing values in the data. Count the number of missing values in each columns.
```{r}
sum(is.na(titanic))
knitr::kable(colSums(is.na(titanic)))
```

#### 10. Calculate the average Age of the passengers. You may want to use the parameter na.rm = TRUE in the function mean.
```{r}
a <- mean(titanic$Age, na.rm = T)
a
```

#### 11. Replace the missing values of age by the average age calculated previously.
```{r}
titanic[is.na(titanic)] <- a
```

#### 12. Remove columns Name, PassengerID, Ticket, and Cabin.
```{r}
titanic_dropped = subset(titanic, select = -c(Name, PassengerId, Ticket, Cabin))
head(titanic_dropped,10)
```

#### 13. Calculate the mean age of female passengers
```{r}
mean(titanic_dropped$Age[titanic_dropped$Sex == "female"])
```

#### 14. Calculate the median fare of the passengers in Class 1
```{r}
median(titanic_dropped$Fare[titanic_dropped$Pclass == 1])
```

#### 15. Calculate the median fare of the female passengers that are not in Class 1
```{r}
median(titanic_dropped$Fare[titanic_dropped$Sex == "female" & titanic_dropped$Pclass != 1])
```

#### 16. Calculate the median age of survived passengers who are female and Class 1 or Class 2
```{r}
median(titanic_dropped$Age[titanic_dropped$Sex == "female" & titanic_dropped$Pclass == 1 | 2])
```

#### 17. Calculate the mean fare of female teenagers survived passengers
```{r}
mean(titanic_dropped$Fare[titanic_dropped$Sex == "female" & titanic_dropped$Survived == 1])

```

#### 18. Calculate the mean fare of female teenagers survived passengers for each class
```{r}
mean(titanic_dropped$Fare[titanic_dropped$Sex == "female" & titanic_dropped$Survived == 1 & titanic_dropped$Pclass == 1 ])

mean(titanic_dropped$Fare[titanic_dropped$Sex == "female" & titanic_dropped$Survived == 1 & titanic_dropped$Pclass == 2 ])

mean(titanic_dropped$Fare[titanic_dropped$Sex == "female" & titanic_dropped$Survived == 1 & titanic_dropped$Pclass == 3 ])

```

#### 19. Calculate the ratio of Survived and not Survived for passengers who are who pays more than the average fare
```{r}
sum(titanic_dropped$Survived[titanic_dropped$Fare > mean(titanic_dropped$Fare)]) / length(titanic_dropped$Survived[titanic_dropped$Fare > mean(titanic_dropped$Fare)])
```

#### 20. Add column that standardizes the fare (subtract the mean and divide by standard deviation) and name it sfare
```{r}
m <- mean(titanic_dropped$Fare)
s <- sd(titanic_dropped$Fare)

sfare <- c((titanic_dropped$Fare - m) / s)
new_Titanic <- cbind(titanic_dropped, sfare)
head(new_Titanic,10)
```

#### 21. Add categorical variable named cfare that takes value cheap for passengers paying less the average fare and takes value expensive for passengers paying more than the average fare.
```{r}
bins <- c(0, mean(new_Titanic$Fare), Inf)
bin_names <- c("cheap", "expensive")
new_Titanic$cfare <- cut(new_Titanic$Fare, breaks = bins, labels = bin_names)
head(new_Titanic,10)
```

#### 22. Add categorical variable named cage that takes value 0 for age 0-10, 1 for age 10-20, 2 for age 20-30, and so on
```{r}
age_bins <- c(0,10,20,30,40,50,60,70,80,90,Inf)
age_bin_names <- c(0,1,2,3,4,5,6,7,8,9)
new_Titanic$cAge <- cut(new_Titanic$Age, breaks = age_bins, labels = age_bin_names)
head(new_Titanic,10)

```

#### 23. Show the frequency of Ports of Embarkation. It appears that there are two missing values in the Embarked variable. Assign the most frequent port to the missing ports. Hint: Use the levels function to modify the categories of categorical variables.
```{r}
table(new_Titanic$Embarked)
levels(new_Titanic$Embarked)[1] <- "S"
table(new_Titanic$Embarked)
```


